Data Encryption Standard Implementation 4003-482-01 RIT

### 2010-05-06 paper/paper.tex
Expanded Software Design section

\section{Software Design}
We chose to implement DES completely in $C++$ for this project.
Both of the members our group were knowledgeable and comfortable
with $C++$ and in the $C$ world in general. We felt we would have a more
generally usable and useful implementation if we went with $C++$ over
another language (Java, Python, ... ). We attempted to take advantage of ac{OOP} but
were careful not to over do it and dive into a downword spiral of subclasses and
analysis of class cohesion. Our $DES$ implementation is simply a class, with actions of encrypting and
decrypting. All state information is organized within the class. We then
implemented a $BlockMode$ class which implements a block stream cipher, which is
just a wrapper around $DES$. The user is then provided with
two small utility functions $encrypt$ \and $decrypt$.

>>>>

\section{Software Design}
Our implementation of DES was done entirely in C++. This was chosen since both members were knowledgeable and comfortable using C++.

Our DES implementation is contained in one class. The implementation is spread across 4 functions; algorithm(), keyschedule(), mixer(), f(). Algorithm() is the starting point. It calls keyschedule() to set up the keys, uses mixer() to perform the initial permutation, then runs the 16 rounds, and ends by using mixer() again to do the inverse initial permutation.  Each round first picks the round's key, calls the f() function, and then swaps the two halves of the key.

The keyschedule() function first uses mixer() to do permutation choice 1.  It then generates 16 sub keys.  Each sub key is generated by shifting the bits 1 or 2 times then using mixer() to do permutation choice 2.

The mixer() function isn't specifically part of the DES spec. DES has many instances where bits must be reordered, we wrote one function, mixer(), which can be passed data and an array defining how to rearrange things and it will rearrange them.

The last function, f(), also known as the Feistel function, is the heart of the encryption.  It first uses mixer() to do the expansion permutation, then xor's the result with the sub key. This output is split into 8-bit block and each block is sent through an S-Box. The output from the S-Boxes is then concatenated run through mixer() using the primitive permutation.
\\ \\ \\
We chose to implement DES completely in $C++$ for this project.
Both of the members our group were knowledgeable and comfortable
with $C++$ and in the $C$ world in general. We felt we would have a more
generally usable and useful implementation if we went with $C++$ over
another language (Java, Python, ... ). We attempted to take advantage of ac{OOP} but
were careful not to over do it and dive into a downword spiral of subclasses and
analysis of class cohesion. Our $DES$ implementation is simply a class, with actions of encrypting and
decrypting. All state information is organized within the class. We then
implemented a $BlockMode$ class which implements a block stream cipher, which is
just a wrapper around $DES$. The user is then provided with
two small utility functions $encrypt$ \and $decrypt$.
